Is Information Ontological or Physical, or Is It Perhaps Something Else? Some Remarks on Krzanowski's Approach to the Concept of Information

### 2023-03-09 ART_Dodig-Crnkovic/Dodig-Crnkovic.tex
Additional corrections added

\begin{artengenv}{Gordana Dodig-Crnkovic}
	{\label{gordana_start}In search of a~common, information-processing, agency-based framework for anthropogenic, biogenic, and abiotic cognition and intelligence}
	{In search of a~common, information-processing, agency-based\ldots}
	{In search of a~common, information-processing, agency-based\\framework for anthropogenic, biogenic, and abiotic cognition\\and intelligence}
	{Chalmers University of Technology,\\University of Gothenburg\\and Mälardalen University, Sweden}
	{Learning from contemporary natural, formal, and social sciences, especially from current biology, as well as from humanities, particularly contemporary philosophy of nature, requires updates of our old definitions of cognition and intelligence. The result of current insights into basal cognition of single cells and evolution of multicellular cognitive systems within the framework of extended evolutionary synthesis (EES) helps us better to understand mechanisms of cognition and intelligence as they appear in nature. New understanding of information and processes of physical (morphological) computation contribute to novel possibilities that can be used to inspire the development of abiotic cognitive systems (cognitive robotics), cognitive computing and artificial intelligence.
	}
	{information, computation, cognition, intelligence, extended evolutionary synthesis, anthropogenic, biogenic and abiotic cognition.}

>>>>

\begin{artengenv}{Gordana Dodig-Crnkovic}
	{\label{gordana_start}In search of a~common, information-processing, agency-based framework for anthropogenic, biogenic, and abiotic cognition and intelligence}
	{In search of a~common, information-processing, agency-based\ldots}
	{In search of a~common, information-processing, agency-based\\framework for anthropogenic, biogenic, and abiotic cognition\\and intelligence}
	{Chalmers University of Technology,\\University of Gothenburg\\and Mälardalen University, Sweden}
	{Learning from contemporary natural, formal, and social sciences, especially from
%	current
	biology, as well as from humanities, particularly contemporary philosophy of nature, requires updates of our old definitions of cognition and intelligence. The result of current insights into basal cognition of single cells and evolution of multicellular cognitive systems within the framework of extended evolutionary synthesis (EES) helps us better to understand mechanisms of cognition and intelligence as they appear in nature. New understanding of information and processes of physical (morphological) computation contribute to novel possibilities that can be used to inspire the development of abiotic cognitive systems (cognitive robotics), cognitive computing and artificial intelligence.
	}
	{information, computation, cognition, intelligence, extended evolutionary synthesis, anthropogenic, biogenic and abiotic cognition.}

### 2023-03-09 ART_Krzanowski_Polak/Krzanowski_Polak.tex
Additional corrections added

\begin{artengenv2auth}{Roman Krzanowski, Pawel Polak}
	{The meta-ontology of AI systems with human-level intelligence}
	{The meta-ontology of AI systems with human-level intelligence}
	{The meta-ontology of AI systems with human-level intelligence}
	{\textsuperscript{1}Departments of Philosophy and Mathematics, Carnegie Mellon University\\
		\textsuperscript{2}Copernicus Center for Interdisciplinary Studies, Jagiellonian University}
	{\label{krzanowpolak_start}In this paper, we examine the meta-ontology of AI systems with human-level intelligence, with us denoting such AI systems as AI\textsubscript{E}. Meta-ontology in philosophy is a~discourse centered on ontology, ontological commitment, and the truth condition of ontological theories. We therefore discuss how meta-ontology is conceptualized for AI\textsubscript{E} systems. We posit that the meta-ontology of AI\textsubscript{E~}systems is not concerned with computational representations of reality in the form of structures, data constructs, or computational concepts, while the ontological commitment of AI\textsubscript{E} systems is directed toward what exists in the outside world. Furthermore, the truth condition of the ontology (which is meta-ontological assumption) of AI\textsubscript{E} systems does not require consistency with closed conceptual schema or ontological theories but rather with reality, or in other words, ``what is the world''
	%\label{ref:RNDuTcZpb4G1k}(Smith, 2019, p.57).
	\parencite[][p.57]{smith_promise_2019}. %
	 In addition, the truth condition of AI\textsubscript{E} systems is verified through operational success rather than by coherence with theories. This work builds on ontological postulates about AI systems that were formulated by Brian Caldwell Smith 
	%\label{ref:RND1W3b57kSJx}(2019).
	\parencite*[][]{smith_promise_2019}.%
	}
	{human-level intelligence AI, meta-ontology of AI Paradigm, AI Paradigm, ontology of AI Paradigm, ontological commitment of AI Paradigm, John Haugeland, Brian Cantwell Smith, Marvin Minsky, Hubert Dreyfus.}
	{%
%		{\flushright\subbold{Roman Krzanowski}\\\subsubsectit\small{Pontifical University of John Paul II in Krakow}\par}%
%		{\flushright\subbold{Pawel Polak}\\\subsubsectit\small{Pontifical University of John Paul II in Krakow}\par}%
		{\flushright\subbold{Roman Krzanowski}\\\subbold{Pawel Polak}\\\subsubsectit\small{Pontifical University of John Paul II in Kraków}\par}%
	}

@@

The ontology of AI\textsubscript{E} systems, meanwhile, denotes and determines how an AI\textsubscript{E} system represents and reasons about the real world. Itis not concerned with computational representations of reality through structures, theories, data constructs, or computational concepts but rather with how real world objects, properties, and relations are registered by an AI system, as well as how they are recognized and interpreted. In brief, this ontology is solely committed to the real world
%\label{ref:RNDHoOzRM31HN}(in the sense explained by Smith, 2019, p.145),
\parencite[in the sense explained by][p.145]{smith_promise_2019}, %
 with it reflecting the real world\footnote{The term ``real world'', or reality, is understood here as it is understood in Smith 
%\label{ref:RNDdF84GECegW}(2019, pp.xiii, xiv),
\parencite*[][xiv]{smith_promise_2019}, %
 denoting the physical world we live in. Minsky 
%\label{ref:RND166uXGx81U}(1991, p.6)
\parencite*[][p.6]{minsky_logical_1991} %
 refers to this reality as common sense reality. The term may be opposed to ``virtual worlds'', ``imaginary worlds'', ``fantastic worlds'', or other qualified uses of ``words'' denoting worlds as creations of computer systems, artistic expressions, or imaginations. The term ``world'' or ``real world'' may have multiple interpretations that we have no intention to discuss as such a~discussion would be pointless and would not further the main point of the paper. Thus, the reader seeking more detailed explanation of this term should follow the cited references..}, or physical reality, and the AI system's place in this world. More specifically, it is embodied, embedded, extended, and enactive 
%\label{ref:RNDcx3KAaKj2v}(Smith, 2019, p.43),
\parencite[][p.43]{smith_promise_2019}, %
 which are terms that will be explained later in this paper. There is no formal theory to accompany 4E ontology (ontology of AI system that is embodied, embedded, extended, and enactive), so there are no criteria for theoretical truth verification, but verification comes instead from a~confrontation with the real world, which we will discuss later. A~4E ontology is not given in the form of a~set of a~priori relations and objects but rather acquired 
%\label{ref:RNDIdA3C5UrQl}(Smith, 2019)
\parencite[][]{smith_promise_2019} %
 in response to a~dynamically changing reality 
%\label{ref:RNDjzZunSO8Xh}(see Minsky, 1991, as well as; Bołtuć, 2020).
\parencites[see][as well as]{minsky_logical_1991}[][]{boltuc_conscious_2020}. %
 We will refer to this ontology as O\textsubscript{E}.

@@

Thus, ontology that represents the real world being always expressed in some form of language (which always is) is always incomplete with respect to the world, i.e., the world ``as is'' cannot be represented by something else, despite the close isomorphism between the reality and some logical aka ontological systems \footnote{All pure ontological systems are logical systems
%\label{ref:RNDQXsIfZlLYA}(Jacquette, 2002, p.xiii; see e.g. Foschini, 2013).
\parencites[][p.xiii]{jacquette_ontology_2002}[see e.g.,][]{foschini_where_2013}.%
} As a~result, there will always be some inconsistency between what exists and what a~given ontology represents, because ontology (even ontology \textbf{O}) will always be a~theory (of sorts), expressed in some specialized language (see ft. 8 ) about the world rather than the world itself. Thus, there will always be a~degree of incommensurability,\footnote{Incommensurability is not understood here in the same way as the incommensurability of paradigms or scientific theory in the works of Khun 
%\label{ref:RNDg7lp3HNe18}(1962)
\parencite*[][]{kuhn_structure_1962} %
 or Feyerabend 
%\label{ref:RNDdDGQ3TRwiM}(e.g. Ryan, 2002; Sankey, 1993; Oberheim and Hoyningen-Huene, 2018; Bird, 2000).
\parencites[e.g.,][]{ryan_feyerabend_2002}[for the incommensurability of paradigms, see also][]{sankey_kuhns_1993}[][]{oberheim_incommensurability_2018}[][]{bird_thomas_2000}. %
 Instead, incommensurability is understood here as a~general sense of not being entirely comparable according to some criteria.} or a~gap, between the ontology of synthetic systems and the reality of what exists. This gap may be narrowed but never entirely closed. Indeed, a~representation can never contain itself as a~part of what exists. We should keep this incommensurability in mind when building synthetic systems like~AI\textsubscript{E}.

@@

Moreover, ontological commitment also includes verifying the criteria for this ontological commitment (i.e., verifying the truth of its existential claims). Ontological commitment can be reduced to the simple claim that ``A man is committed to the truth of whatever he asserts''
%\label{ref:RNDA5OgzdmbIa}(Searle, 1969, p.112).
\parencite[][p.112]{searle_speech_1969}.%
\footnote{See the critique of Searle by 
%\label{ref:RNDdyJS6eDQgJ}(Inwagen, 1991).
\parencite[][]{lepore_searle_1991}.%
} For some, however, particularly professional ontologists, Searle's take on ontological commitment seems perfunctory, but in the AI context, it provides a~simple (i.e., operational) means for judging the scope of an AI ontology. We do not dispute the robustness of Searle's claim on ontological commitment; here we take it as a~guide in practical applications.

Truth conditions are what makes an ontology correct, such that a~``theory is ontologically committed to an object only if that object occurs in all the ontologies of that theory''
%\label{ref:RND7ceF0NY8f3}(Gibson, 2009, p.631).
\parencite[][p.631]{gibson_ontological_2009}. %
 This statement was rephrased by Rayo 
%\label{ref:RND2olhyVRiGx}(2007)
\parencite*[][]{rayo_ontological_2007} %
 into the following claim: ``…for a~sentence to carry commitment to Fs is for the sentence's truth to demand of the world that it contain Fs.'' Gibson's claim therefore comes from a~philosophical perspective. It requires that a~theory (of ontology) can be expressed in sentences (of any language), so this truth condition boils down to an agreement between the theory and the world in question (the correspondence theory of truth\footnote{We are not going here into the discussion of the correspondence theory of truth or truth in general. We assume that for the systems (natural or artificial) ``being in the world'' (no Heideggerian connotations) there must be some relation, however tentative and limited holding between them and the real world they are immersed in, the relation of truth. This relation in a~case of evolution and synthetic systems is verified by systems' operational success 
%\label{ref:RND1clmQWWHtj}(see for example the discussion of truth relation in Bird, 2000).
\parencite[see for example the discussion of truth relation in][]{bird_thomas_2000}. %
 The correspondence theory of truth seems to trouble only philosophers of the anti-realist, skeptical persuasion, not computer scientists or evolutionary biologists 
%\label{ref:RNDL8KkhQlfH3}(Bird, 2000).
\parencite[][]{bird_thomas_2000}. %
 ``Operational success'' in applied sciences may be thought as playing the role of empirical proof in naturalized epistemology 
%\label{ref:RNDDKaQHD0jpK}(Bird, 2000, p.263).
\parencite[][p.263]{bird_thomas_2000}. %
 }), which may be the real world (as is the case of AI\textsubscript{E} systems) or an imaginary constructs or virtual realities whatever the domain of ontology is.

>>>>

\begin{artengenv2auth}{Roman Krzanowski, Pawel Polak}
	{The meta-ontology of AI systems with human-level intelligence}
	{The meta-ontology of AI systems with human-level intelligence}
	{The meta-ontology of AI systems with human-level intelligence}
	{\textsuperscript{1}Departments of Philosophy and Mathematics, Carnegie Mellon University\\
		\textsuperscript{2}Copernicus Center for Interdisciplinary Studies, Jagiellonian University}
	{\label{krzanowpolak_start}In this paper, we examine the meta-ontology of AI systems with human-level intelligence, with us denoting such AI systems as AI\textsubscript{E}. Meta-ontology in philosophy is a~discourse centered on ontology, ontological commitment, and the truth condition of ontological theories. We therefore discuss how meta-ontology is conceptualized for AI\textsubscript{E} systems. We posit that the meta-ontology of AI\textsubscript{E~}systems is not concerned with computational representations of reality in the form of structures, data constructs, or computational concepts, while the ontological commitment of AI\textsubscript{E} systems is directed toward what exists in the outside world. Furthermore, the truth condition of the ontology (which is meta-ontological assumption) of AI\textsubscript{E} systems does not require consistency with closed conceptual schema or ontological theories but rather with reality, or in other words, ``what is the world''
	%\label{ref:RNDuTcZpb4G1k}(Smith, 2019, p.57).
	\parencite[][p.57]{smith_promise_2019}. %
	 In addition, the truth condition of AI\textsubscript{E} systems is verified through operational success rather than by coherence with theories. This work builds on ontological postulates about AI systems that were formulated by Brian Cantwell Smith 
	%\label{ref:RND1W3b57kSJx}(2019).
	\parencite*[][]{smith_promise_2019}.%
	}
	{human-level intelligence AI, meta-ontology of AI Paradigm, AI Paradigm, ontology of AI Paradigm, ontological commitment of AI Paradigm, John Haugeland, Brian Cantwell Smith, Marvin Minsky, Hubert Dreyfus.}
	{%
%		{\flushright\subbold{Roman Krzanowski}\\\subsubsectit\small{Pontifical University of John Paul II in Krakow}\par}%
%		{\flushright\subbold{Pawel Polak}\\\subsubsectit\small{Pontifical University of John Paul II in Krakow}\par}%
		{\flushright\subbold{Roman Krzanowski}\\\subbold{Pawel Polak}\\\subsubsectit\small{Pontifical University of John Paul II in Kraków, Poland}\par}%
	}

@@

The ontology of AI\textsubscript{E} systems, meanwhile, denotes and determines how an AI\textsubscript{E} system represents and reasons about the real world. Itis not concerned with computational representations of reality through structures, theories, data constructs, or computational concepts but rather with how real world objects, properties, and relations are registered by an AI system, as well as how they are recognized and interpreted. In brief, this ontology is solely committed to the real world
%\label{ref:RNDHoOzRM31HN}(in the sense explained by Smith, 2019, p.145),
\parencite[in the sense explained by][p.145]{smith_promise_2019}, %
 with it reflecting the real world\footnote{The term ``real world'', or reality, is understood here as it is understood in Smith 
%\label{ref:RNDdF84GECegW}(2019, pp.xiii, xiv),
\parencite*[][xiv]{smith_promise_2019}, %
 denoting the physical world we live in. Minsky 
%\label{ref:RND166uXGx81U}(1991, p.6)
\parencite*[][p.6]{minsky_logical_1991} %
 refers to this reality as common sense reality. The term may be opposed to ``virtual worlds'', ``imaginary worlds'', ``fantastic worlds'', or other qualified uses of ``words'' denoting worlds as creations of computer systems, artistic expressions, or imaginations. The term ``world'' or ``real world'' may have multiple interpretations that we have no intention to discuss as such a~discussion would be pointless and would not further the main point of the paper. Thus, the reader seeking more detailed explanation of this term should follow the cited references.}, or physical reality, and the AI system's place in this world. More specifically, it is embodied, embedded, extended, and enactive 
%\label{ref:RNDcx3KAaKj2v}(Smith, 2019, p.43),
\parencite[][p.43]{smith_promise_2019}, %
 which are terms that will be explained later in this paper. There is no formal theory to accompany 4E ontology (ontology of AI system that is embodied, embedded, extended, and enactive), so there are no criteria for theoretical truth verification, but verification comes instead from a~confrontation with the real world, which we will discuss later. A~4E ontology is not given in the form of a~set of a~priori relations and objects but rather acquired 
%\label{ref:RNDIdA3C5UrQl}(Smith, 2019)
\parencite[][]{smith_promise_2019} %
 in response to a~dynamically changing reality 
%\label{ref:RNDjzZunSO8Xh}(see Minsky, 1991, as well as; Bołtuć, 2020).
\parencites[see][as well as]{minsky_logical_1991}[][]{boltuc_conscious_2020}. %
 We will refer to this ontology as O\textsubscript{E}.

@@

Thus, ontology that represents the real world being always expressed in some form of language (which always is) is always incomplete with respect to the world, i.e., the world ``as is'' cannot be represented by something else, despite the close isomorphism between the reality and some logical aka ontological systems.\footnote{All pure ontological systems are logical systems
%\label{ref:RNDQXsIfZlLYA}(Jacquette, 2002, p.xiii; see e.g. Foschini, 2013).
\parencites[][p.xiii]{jacquette_ontology_2002}[see e.g.,][]{foschini_where_2013}.%
} As a~result, there will always be some inconsistency between what exists and what a~given ontology represents, because ontology (even ontology \textbf{O}) will always be a~theory (of sorts), expressed in some specialized language (see ft. 8) about the world rather than the world itself. Thus, there will always be a~degree of incommensurability,\footnote{Incommensurability is not understood here in the same way as the incommensurability of paradigms or scientific theory in the works of Khun 
%\label{ref:RNDg7lp3HNe18}(1962)
\parencite*[][]{kuhn_structure_1962} %
 or Feyerabend 
%\label{ref:RNDdDGQ3TRwiM}(e.g. Ryan, 2002; Sankey, 1993; Oberheim and Hoyningen-Huene, 2018; Bird, 2000).
\parencites[e.g.,][]{ryan_feyerabend_2002}[for the incommensurability of paradigms, see also][]{sankey_kuhns_1993}[][]{oberheim_incommensurability_2018}[][]{bird_thomas_2000}. %
 Instead, incommensurability is understood here as a~general sense of not being entirely comparable according to some criteria.} or a~gap, between the ontology of synthetic systems and the reality of what exists. This gap may be narrowed but never entirely closed. Indeed, a~representation can never contain itself as a~part of what exists. We should keep this incommensurability in mind when building synthetic systems like~AI\textsubscript{E}.

@@

Moreover, ontological commitment also includes verifying the criteria for this ontological commitment (i.e., verifying the truth of its existential claims). Ontological commitment can be reduced to the simple claim that ``A man is committed to the truth of whatever he asserts''
%\label{ref:RNDA5OgzdmbIa}(Searle, 1969, p.112).
\parencite[][p.112]{searle_speech_1969}.%
\footnote{See the critique of Searle by Inwagen
%\label{ref:RNDdyJS6eDQgJ}(Inwagen, 1991).
\parencite*[][]{lepore_searle_1991}.%
} For some, however, particularly professional ontologists, Searle's take on ontological commitment seems perfunctory, but in the AI context, it provides a~simple (i.e., operational) means for judging the scope of an AI ontology. We do not dispute the robustness of Searle's claim on ontological commitment; here we take it as a~guide in practical applications.

Truth conditions are what makes an ontology correct, such that a~``theory is ontologically committed to an object only if that object occurs in all the ontologies of that theory''
%\label{ref:RND7ceF0NY8f3}(Gibson, 2009, p.631).
\parencite[][p.631]{gibson_ontological_2009}. %
 This statement was rephrased by Rayo 
%\label{ref:RND2olhyVRiGx}(2007)
\parencite*[][]{rayo_ontological_2007} %
 into the following claim: ``[…] for a~sentence to carry commitment to Fs is for the sentence's truth to demand of the world that it contain Fs.'' Gibson's claim therefore comes from a~philosophical perspective. It requires that a~theory (of ontology) can be expressed in sentences (of any language), so this truth condition boils down to an agreement between the theory and the world in question (the correspondence theory of truth\footnote{We are not going here into the discussion of the correspondence theory of truth or truth in general. We assume that for the systems (natural or artificial) ``being in the world'' (no Heideggerian connotations) there must be some relation, however tentative and limited holding between them and the real world they are immersed in, the relation of truth. This relation in a~case of evolution and synthetic systems is verified by systems' operational success 
%\label{ref:RND1clmQWWHtj}(see for example the discussion of truth relation in Bird, 2000).
\parencite[see for example the discussion of truth relation in][]{bird_thomas_2000}. %
 The correspondence theory of truth seems to trouble only philosophers of the anti-realist, skeptical persuasion, not computer scientists or evolutionary biologists 
%\label{ref:RNDL8KkhQlfH3}(Bird, 2000).
\parencite[][]{bird_thomas_2000}. %
 ``Operational success'' in applied sciences may be thought as playing the role of empirical proof in naturalized epistemology 
%\label{ref:RNDDKaQHD0jpK}(Bird, 2000, p.263).
\parencite[][p.263]{bird_thomas_2000}. %
 }), which may be the real world (as is the case of AI\textsubscript{E} systems) or an imaginary constructs or virtual realities whatever the domain of ontology is.

### 2023-03-09 ART_Martin/Martin.tex
Additional corrections added

\begin{artengenv2auth}{Alice Martin, Mathieu Magnaudet, Stéphane Conversy}
	{Modelling interactive computing systems: Do we have a~good theory of what computers are?}
	{Modelling interactive computing systems\ldots}
	{Modelling interactive computing systems: Do we have a~good theory\\of what computers are?}
	{\textsuperscript{1}Departments of Philosophy and Mathematics, Carnegie Mellon University\\
		\textsuperscript{2}Copernicus Center for Interdisciplinary Studies, Jagiellonian University}
	{Computers are increasingly interactive. They are no more transformational systems producing a~final output after a~finite execution. Instead, they continuously react in time to external events that modify the course of computing execution. While philosophers have been interested in conceptualizing computers for a~long time, they seem to have paid little attention to the specificities of interactive computing. We propose to tackle this issue by surveying the literature in theoretical computer science, where one can find explicit proposals for a~model of interactive computing. 
	In that field, the formal modelling of interactive computing systems has been brought down to whether the new interaction models are reducible to Turing Machines. There are three areas where interaction models are framed. The comparison between TMs and interactive system models is at stake in all of them. These areas are namely some works on concurrency by Milner, on Reactive Turing Machines, and on interaction as a~new computing paradigm. For each of the three identified models, we present its motivation, sum up its account for interaction and its legacy, and point out issues regarding the understanding of computers. The survey shows difficulties for epistemologists. The reason is that these analyses focus on the formal equivalence between interactive models of computation and classic ones. Such a~project is different from addressing how a~computing machine can be interactive: in other words, which mechanisms allow it.}
	{philosophy of computing, models of computation, interactive computing, computing mechanism, computational mechanistic explanation.}
	{%
%		{\flushright\subbold{Alice Martin}\\\subsubsectit\small{ENAC, Université de Toulouse}\par}%
%		{\flushright\subbold{Mathieu Magnaudet}\\\subsubsectit\small{ENAC, Université de Toulouse}\par}%
%		{\flushright\subbold{Stéphane Conversy}\\\subsubsectit\small{ENAC, Université de Toulouse}\par}%
				{\flushright\subbold{Alice Martin}\\\subbold{Mathieu Magnaudet}\\\subbold{Stéphane Conversy}\\
				\subsubsectit\small{ENAC, Université de Toulouse}\par}%
	}

@@

From this, Wegner concludes that computing systems should not be thought of as algorithms but as \textit{interfaces}, \textit{views}, and \textit{modes of use}, definable as behaviours to be specified. Consequently, an ontological question is also at stake: in what terms should the external world be modelled: as atomic objects and events? As processes and flow?
Formally, Wegner’s account of interaction has led to the development of Persistent Turing machines (PTMs), a~model of sequential computation, and the result that multi-stream interaction machines (MIMs) are more expressive than sequential interaction machines (SIMs) \parencite{Goldin2000, Goldin2004}. Wegner and Goldin trace back the idea that interaction is not expressible by or reducible to algorithms at the closing conference on the 5\textsuperscript{th}-generation computing project in the context of logic programming. Reactiveness of logic programs, realized by the commitment to a~course of action, was shown to be incompatible with logical completeness \parencite{Wegner1999}.

@@

So, do we have a~good theory about interactive computers? Do we understand what they are? A~natural and common way to go is to reduce the question of what interactive computers are to what interactive computation is. Initially, the first models of computation emerged through computability theory. They served as answers to an abstract mathematical problem, namely the formalization of the intuitive notion of an algorithm. They had nothing to say about computers, as computers did not even exist at the time. Since the computability era, models of computation like the Turing Machines have been exported outside their original scope to serve as a~basis for theoretical computer science. Some models of computation (Turing Machines) have even helped to reflect on computers. It is no surprise since computers were thought to be precisely the kind of machines that implement computations. Models of computation have then evolved, accounting for new desired properties to be integrated within the classical framework. In computer science, what makes a~model of computation valuable is related to the formal properties it expresses. Once those formal properties are at hand, they allow further procedures to be acted upon them, especially system verification and certification.  In the end, models of computation serve as tools used to support and verify a~system’s design. These models belong to a~particular abstraction level: they do not intend to model the system as a~whole and the way it works. They focus on verifiable properties, upon which proofs that guarantee the outputs of the system are built. Verifying formal properties is different from investigating why the system behaves the way it does. They are two different tasks. The former task (verification) belongs to applied mathematics. It describes abstract computations through formal models by focusing on specific properties. The latter (understanding computing behaviour) is the question the philosopher begs when asking what a~computer is. It requires something else than task-oriented formalizations of properties abstracted away from any physical mechanism. Philosophers of computing need to make sense of the overall behaviour, which requires combining other levels of abstraction.  The reason is that an account of computing behaviour calls upon the description of how computation can be carried out: in other words, it requires the description of execution on some computer architecture.
Computations and their models belong to a~level of abstraction independent from implementation detail. Computations, as already coined, are ``medium-independent'' \parencite{Klein2020}. On the contrary, to have a~model of some execution belongs to a~lower level of abstraction, where minimal references to the devices that allow the execution is made. There is no need to dig into fine-grained implementation details to make sense of computing behaviour in mechanistic terms.

>>>>

\begin{artengenv2auth}{Alice Martin, Mathieu Magnaudet, Stéphane Conversy}
	{Modelling interactive computing systems: Do we have a~good theory of what computers are?}
	{Modelling interactive computing systems\ldots}
	{Modelling interactive computing systems: Do we have a~good theory\\of what computers are?}
	{\textsuperscript{1}Departments of Philosophy and Mathematics, Carnegie Mellon University\\
		\textsuperscript{2}Copernicus Center for Interdisciplinary Studies, Jagiellonian University}
	{Computers are increasingly interactive. They are no more transformational systems producing a~final output after a~finite execution. Instead, they continuously react in time to external events that modify the course of computing execution. While philosophers have been interested in conceptualizing computers for a~long time, they seem to have paid little attention to the specificities of interactive computing. We propose to tackle this issue by surveying the literature in theoretical computer science, where one can find explicit proposals for a~model of interactive computing. 
	In that field, the formal modelling of interactive computing systems has been brought down to whether the new interaction models are reducible to Turing Machines. There are three areas where interaction models are framed. The comparison between TMs and interactive system models is at stake in all of them. These areas are namely some works on concurrency by Milner, on Reactive Turing Machines, and on interaction as a~new computing paradigm. For each of the three identified models, we present its motivation, sum up its account for interaction and its legacy, and point out issues regarding the understanding of computers. The survey shows difficulties for epistemologists. The reason is that these analyses focus on the formal equivalence between interactive models of computation and classic ones. Such a~project is different from addressing how a~computing machine can be interactive: in other words, which mechanisms allow it.}
	{philosophy of computing, models of computation, interactive computing, computing mechanism, computational mechanistic explanation.}
	{%
%		{\flushright\subbold{Alice Martin}\\\subsubsectit\small{ENAC, Université de Toulouse}\par}%
%		{\flushright\subbold{Mathieu Magnaudet}\\\subsubsectit\small{ENAC, Université de Toulouse}\par}%
%		{\flushright\subbold{Stéphane Conversy}\\\subsubsectit\small{ENAC, Université de Toulouse}\par}%
				{\flushright\subbold{Alice Martin}\\\subbold{Mathieu Magnaudet}\\\subbold{Stéphane Conversy}\\
				\subsubsectit\small{ENAC, Université de Toulouse, France}\par}%
	}

@@

From this, Wegner concludes that computing systems should not be thought of as algorithms but as \textit{interfaces}, \textit{views}, and \textit{modes of use}, definable as behaviours to be specified. Consequently, an ontological question is also at stake: in what terms should the external world be modelled: as atomic objects and events? As processes and flow?
Formally, Wegner’s account of interaction has led to the development of Persistent Turing machines (PTMs), a~model of sequential computation, and the result that multi-stream interaction machines (MIMs) are more expressive than sequential interaction machines (SIMs) \parencite{Goldin2000, Goldin2004}. Wegner and Goldin trace back the idea that interaction is not expressible by or reducible to algorithms at the closing conference on the 5\textsuperscript{th}-Generation Computer Project in the context of logic programming. Reactiveness of logic programs, realized by the commitment to a~course of action, was shown to be incompatible with logical completeness \parencite{Wegner1999}.

@@

So, do we have a~good theory about interactive computers? Do we understand what they are? A~natural and common way to go is to reduce the question of what interactive computers are to what interactive computation is. Initially, the first models of computation emerged through computability theory. They served as answers to an abstract mathematical problem, namely the formalization of the intuitive notion of an algorithm. They had nothing to say about computers, as computers did not even exist at the time. Since the computability era, models of computation like the Turing Machines have been exported outside their original scope to serve as a~basis for theoretical computer science. Some models of computation (Turing Machines) have even helped to reflect on computers. It is no surprise since computers were thought to be precisely the kind of machines that implement computations. Models of computation have then evolved, accounting for new desired properties to be integrated within the classical framework. In computer science, what makes a~model of computation valuable is related to the formal properties it expresses. Once those formal properties are at hand, they allow further procedures to be acted upon them, especially system verification and certification.  In the end, models of computation serve as tools used to support and verify a~system’s design. These models belong to a~particular abstraction level: they do not intend to model the system as a~whole and the way it works. They focus on verifiable properties, upon which proofs that guarantee the outputs of the system are built. Verifying formal properties is different from investigating why the system behaves the way it does. They are two different tasks. The former task (verification) belongs to applied mathematics. It describes abstract computations through formal models by focusing on specific properties. The latter (understanding computing behaviour) is the question the philosopher begs when asking what a~computer is. It requires something else than task-oriented formalizations of properties abstracted away from any physical mechanism. Philosophers of computing need to make sense of the overall behaviour, which requires combining other levels of abstraction.  The reason is that an account of computing behaviour calls upon the description of how computation can be carried out: in other words, it requires the description of execution on some computer architecture.
Computations and their models belong to a~level of abstraction independent from implementation detail. Computations, as already coined, are ``medium-independent'' \parencite{Klein2020}. On the contrary, to have a~model of some execution belongs to a~lower level of abstraction, where minimal references to the devices that allow the execution are made. There is no need to dig into fine-grained implementation details to make sense of computing behaviour in mechanistic terms.

### 2023-03-09 ART_Mscislawski/Mscislaw.tex
Additional corrections added

\begin{artengenv}{Łukasz Mścisławski}
	{Is information ontological or physical, or is it perhaps something else? Some remarks on Krzanowski's approach to the concept of information}
	{Is information ontological or physical, or is it perhaps\ldots}
	{Is information ontological or physical, or is it perhaps something\\else? Some remarks on Krzanowski's approach to the concept of\\information}
	{Wrocław University of Science and Technology}
	{\label{mscislaw_start}As one may have noticed, the title of this paper is somewhat provocative. We found Roman Krzanowski's
	%\label{ref:RNDD4ZPbWpWxt}(2020a; 2020b; 2020c; 2022)
	\parencites*[][]{krzanowski_does_2020}[][]{krzanowski_what_2020}[][]{krzanowski_why_2020}[][]{krzanowski_ontological_2022} %
	 proposed approach to the problem of information very intriguing. Our aim here is to highlight some advantages when it comes to answering some fundamental questions in the philosophy of physics and metaphysics, as well as the philosophy of information and computer science. This issue is of great importance, so we propose that the introduction of some subtle distinctions between ontological and epistemological information can be regarded as being analogous to G.F.R. Ellis's analyses of the passage of time in his concept of the Crystallizing Block Universe 
	%\label{ref:RNDaS7kEBPadF}(Ellis and Goswami, 2012).
	\parencite[][]{ellis_space_2012}. %
	 This analogy could be useful when further studying the relations between different types of information. We also suggest some subjects for further study, ones where Krzanowski's proposal could serve as a~very solid foundation for examining traditional metaphysical issues by combining classical philosophical doctrines with the new approach.
	}
	{physical information, ontology, physics, philosophy of information.}

>>>>

\begin{artengenv}{Łukasz Mścisławski}
	{Is information ontological or physical, or is it perhaps something else? Some remarks on Krzanowski's approach to the concept of information}
	{Is information ontological or physical, or is it perhaps\ldots}
	{Is information ontological or physical, or is it perhaps something\\else? Some remarks on Krzanowski's approach to the concept of\\information}
	{Wrocław University of Science and Technology, Poland}
	{\label{mscislaw_start}As one may have noticed, the title of this paper is somewhat provocative. We found Roman Krzanowski's
	%\label{ref:RNDD4ZPbWpWxt}(2020a; 2020b; 2020c; 2022)
	\parencites*[][]{krzanowski_does_2020}[][]{krzanowski_what_2020}[][]{krzanowski_why_2020}[][]{krzanowski_ontological_2022} %
	 proposed approach to the problem of information very intriguing. Our aim here is to highlight some advantages when it comes to answering some fundamental questions in the philosophy of physics and metaphysics, as well as the philosophy of information and computer science. This issue is of great importance, so we propose that the introduction of some subtle distinctions between ontological and epistemological information can be regarded as being analogous to G.F.R. Ellis's analyses of the passage of time in his concept of the Crystallizing Block Universe 
	%\label{ref:RNDaS7kEBPadF}(Ellis and Goswami, 2012).
	\parencite[][]{ellis_space_2012}. %
	 This analogy could be useful when further studying the relations between different types of information. We also suggest some subjects for further study, ones where Krzanowski's proposal could serve as a~very solid foundation for examining traditional metaphysical issues by combining classical philosophical doctrines with the new approach.
	}
	{physical information, ontology, physics, philosophy of information.}
